opponent modelling

### 2019-11-11 comp_model_1019.Rmd
added self tuned EWA

```{r}
QL_modelling <- list()
for(id in unique(dat$human_id)) {
  QL_modelling[[id]] <- list()
  tdat <- subset(dat,human_id == id)
  QL_modelling[[id]] <- optim(c(0.1,0.1),fn=Q_learn,gr = NULL, data=tdat,"-2loglik", lower = c(0.001,0.01), upper = c(0.99,10), method="L-BFGS-B")
}

save(QL_modelling,file="QL_modelling.RData")

@@

## Q-learning with state space including last actions

>>>>

```{r}
# QL_modelling <- list()
# for(id in unique(dat$human_id)) {
#   QL_modelling[[id]] <- list()
#   tdat <- subset(dat,human_id == id)
#   QL_modelling[[id]] <- optim(c(0.1,0.1),fn=Q_learn,gr = NULL, data=tdat,"-2loglik", lower = c(0.001,0.01), upper = c(0.99,10), method="L-BFGS-B")
# }
# 
# save(QL_modelling,file="QL_modelling.RData")
# 
# 
# ```
# 
# 
# ## Parametric EWA 
# 
# ```{r}
# 
# # Define reward matrices from the prospective of the row player (human in our case)
reward_RPS <- t(matrix(c(0,-1,1,1,0,-1,-1,1,0),nrow=3,ncol=3))
dimnames(reward_RPS) = list(c("R", "P", "S"), c("R", "P", "S"))

reward_FWG <- t(matrix(c(0,-1,1,1,0,-1,-1,1,0),nrow=3,ncol=3))
dimnames(reward_FWG) = list(c("F", "W", "G"), c("F", "W", "G"))

reward_NUM <- t(matrix(c(0,-1,0,0,1,1,0,-1,0,0,0,1,0,-1,0,0,0,1,0,-1,-1,0,0,1,0),nrow=5, ncol=5))
dimnames(reward_NUM) = list(c("1", "2", "3", "4", "5"), c("1", "2", "3", "4", "5"))

# 
# 
# EWA_par <- function(par,data,return_value=c("-2loglik","likelihood_by_trial")){
#   # par: vector of parameters to the EWA model 
#   # Data :choice data by trial for each participant 
#   # Returns sum of loglikelihoods or likelihood per trial 
#   phi <- par[1]
#   delta <- par[2]
#   rho <- par[3]
#   lambda <- par[4]
#   
#   # Initiate N(0) = 1 as in Camerer and Ho 1997 paper. 
#   N <- 1
#   
#   # Define attraction vectors for each game 
#   A_RPS = matrix(0.0,3)
#   names(A_RPS) <- c("R","P","S")
#   A_FWG = matrix(0.0,3)
#   names(A_FWG) <- c("F","W","G")
#   A_NUM = matrix(0.0,5)
#   names(A_NUM) <- c("1","2","3","4","5")
#   
#   # Initiate likelihood by trial vector 
#   lik_hum <- matrix(0.0,nrow(data))
# 
#   for(t in 1:nrow(data)) {
#     t_game <- data[t,"game"]
#     if(t_game == "numbers") nopts <- 5 else nopts <- 3
#     if(data[t,"round"] == 1) {
#       # first round is uniform prediction
#       Att <- switch(as.character(t_game),rps=A_RPS,fwg = A_FWG, numbers = A_NUM)
#       reward <- switch(as.character(t_game),rps=reward_RPS,fwg = reward_FWG, numbers = reward_NUM)
#       lik_hum[t] <- 1/nopts
#     } else {
#       # Get reward and past human action
#       h_act_prev <- as.character(data[t-1,"h_action"])
#       a_act_prev <- as.character(data[t-1,"a_action"])
#       for (i in length(Att)) {
#         action <- as.character(names(Att)[i])
#         # cat("this is current strat",action,"\n")
#         # cat("this is previous humna action",h_act_prev,"\n")
#         
#         # Attraction vector update rule 
#         Att[[i]] <- (phi*N*Att[[i]]  + ( delta + (1- delta)*(action == h_act_prev))*as.numeric(reward[action,a_act_prev])) / (rho*N + 1)
#       }
#       #Update the value of N 
#       N <- rho*N + 1 
#       
#       # Assume human chooses action probabilistically using softmax on Attraction values
#       probs <- exp(lambda*Att)/sum(exp(lambda*Att))
#       
#       # Get actual human action and compute likelihood
#       h_act <- as.character(data[t,"h_action"])
#       act_index <- match(h_act, names(Att))
#       lik_hum[t] <- probs[[act_index]]
#     }
#   }
#   
#  if(return_value == "-2loglik") {
#     ret <- -2*sum(log(lik_hum))
#     if(is.infinite(ret)) {
#       return(1e+300)
#     } else {
#       return(ret)
#     }
#   }
#   if(return_value == "likelihood_by_trial") return(lik_hum)
# }
```

```{r}
#######################################################################################

##  Self-Tuning EWA:

EWA_self <- function(par,data,return_value=c("-2loglik","likelihood_by_trial")){

  lambda <- par[[1]]
  # Initiate N(0) = 1 as in Camerer and Ho 1997 paper. 
  N <- 1.0

  # Define attraction vectors for each game 
  A_RPS = matrix(0.0,3)
  names(A_RPS) <- c("R","P","S")
  A_FWG = matrix(0.0,3)
  names(A_FWG) <- c("F","W","G")
  A_NUM = matrix(0.0,5)
  names(A_NUM) <- c("1","2","3","4","5")

  # Initiate likelihood by trial vector 
  lik_hum <- matrix(0.0,nrow(data))

  for(t in 1:nrow(data)) {
    t_game <- data[t,"game"]
    if(t_game == "numbers") nopts <- 5 else nopts <- 3
    if(data[t,"round"] == 1) {
      # first round is uniform prediction
      Att <- switch(as.character(t_game),rps=A_RPS,fwg = A_FWG, numbers = A_NUM)
      reward <- switch(as.character(t_game),rps=reward_RPS,fwg = reward_FWG, numbers = reward_NUM)
      game_data <- subset(data, game == t_game)
      lik_hum[t] <- 1/nopts

      R_t <- rep(0.0,nopts)
      H_t <- R_t

    } else {

      indx <- data[t,"round"]
      # Get reward and past human action
      h_act_prev <- as.character(game_data[indx-1,"h_action"])
      a_act_prev <- as.character(game_data[indx-1,"a_action"])

      # Estimate phi(t)
      R_t <- as.numeric(names(Att) == a_act_prev)
      # cat("This is R(t)",R_t,"\n")

      H_t <- (H_t*(indx-2) + R_t)/(indx-1)
      # cat("this is H(t)",H_t,"\n")

      Phi_t <- 1 - 0.5*sum((H_t-R_t)^2)
      # cat("This is Phit(t)",Phi_t,"\t")

      # Estimate vector Delta(t)
      delta_t <- as.numeric(reward[,a_act_prev] >= as.numeric(game_data[indx-1,"score"]))
      # cat("this is delta(t)",delta_t,"\n")

      for (i in length(Att)) {
        action <- as.character(names(Att)[i])
        # cat("this is current strat",action,"\n")
        # cat("this is previous humna action",h_act_prev,"\n")

        # Attraction vector update rule 
        Att[[i]] <- (Phi_t*N*Att[[i]]  + ( delta_t[[i]] + (1- delta_t[[i]])*(action == h_act_prev))*as.numeric(reward[action,a_act_prev])) / (Phi_t*N + 1)
      }
      #Update the value of N 
      N <- Phi_t*N + 1

      # Assume human chooses action probabilistically using softmax on Attraction values
      probs <- exp(lambda*Att)/sum(exp(lambda*Att))

      # Get actual human action and compute likelihood
      h_act <- as.character(game_data[indx,"h_action"])
      act_index <- match(h_act, names(Att))
      lik_hum[t] <- probs[[act_index]]
    }
  }

 if(return_value == "-2loglik") {
    ret <- -2*sum(log(lik_hum))
    if(is.infinite(ret)) {
      return(1e+300)
    } else {
      return(ret)
    }
  }
  if(return_value == "likelihood_by_trial") return(lik_hum)
}

@@

```{r}
# 
# EWA_modelling <- list()
# for(id in unique(dat$human_id)) {
# 
#   EWA_modelling[[id]] <- list()
#   tdat <- subset(dat,human_id == id)
#   EWA_modelling[[id]] <- optim(c(0.5,0.5,0.5,0.5),fn=EWA_par,gr = NULL, data=tdat,"-2loglik", lower = c(0.001,0.001,0.001,0.001), upper = c(1,0.99,0.99,1), method="L-BFGS-B")
# }
# 
# save(EWA_modelling,file="EWA_modelling.RData")

EWA_self_modelling <- list()
for(id in unique(dat$human_id)) {
  EWA_self_modelling[[id]] <- list()
  tdat <- subset(dat,human_id == id)
  EWA_self_modelling[[id]] <- optim(10.0,fn=EWA_self,gr = NULL, data=tdat,"-2loglik", lower = 0.001, upper = 200, method="L-BFGS-B")
}

save(EWA_self_modelling,file="EWA_self_modelling.RData")
